---
title: "APPLIED MACHINE LEARNING FINAL PROJECT"
output:
  html_document: default
  pdf_document: default
font-family: Gill Sans
date: "`r format(Sys.time(), '%d %B %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GROUP MEMBERS
Hakkıcan BÜLÜÇ-180709061 / Adem VAROL-200709078 / Hüseyin ÇOBAN-180709018

## DATA SUMMARY   
Data read from excel file
```{r}
#install.packages("readxl")
library(readxl)
# Enter the path where the data is
df <- read_excel("AML_DATA.xls")
names(df)
summary(df$ORTALAMA)
```

```{r}
#install.packages("dplyr")
library(dplyr)

# Student count
sprintf("Number of students in school OAL  : %d",nrow(filter(df, OKUL == "OAL")))
sprintf("Number of students in school YEAL : %d",nrow(filter(df, OKUL == "YEAL")))
```

```{r}
library(data.table)
#CLASSES
dt9 <- data.table(df)[SINIF %like% "9"]
dt10 <- data.table(df)[SINIF %like% "10"]
dt11 <- data.table(df)[SINIF %like% "11"]
dt12 <- data.table(df)[SINIF %like% "12"]

# CLASS COUNT
sprintf(" 9. GRADE: %d",nrow(dt9))
sprintf("10. GRADE: %d",nrow(dt10))
sprintf("11. GRADE: %d",nrow(dt11))
sprintf("12. GRADE: %d",nrow(dt12))
```
```{r}
# GENDER COUNT
sprintf("NUMBER OF GIRLS: %d",nrow(filter(df, CINSIYET == 0)))
sprintf("NUMBER OF BOYS : %d",nrow(filter(df, CINSIYET == 1)))
```

```{r}
# CITY - VILLAGE COUNT
sprintf("Number of people living in the village : %d",nrow(filter(df, IKAMET == 0)))
sprintf("Number of people living in the city    : %d",nrow(filter(df, IKAMET == 1)))
```

```{r}
# MEANS

sprintf("Cumulative grade point average : %f",mean(df$ORTALAMA))
sprintf("rade point average of OAL      : %f",mean(filter(df, OKUL == "OAL")$ORTALAMA))
sprintf("rade point average of YEAL     : %f",mean(filter(df, OKUL == "YEAL")$ORTALAMA))
```

```{r}
# MOST SUCCESFUL STUDENT
sprintf("The student with the highest average : %s",df$`ADI SOYADI`[match(max(df$ORTALAMA), df$ORTALAMA)])
```

```{r}
#Bar graph showing the most elementary school graduations of OAL students' parents
oal <- filter(df, OKUL == "OAL")
oal <- oal[, (names(oal) %in% c("CINSIYET", "Anne en fazla ilkokul mezunu mu?", "Baba en fazla ilkokul mezunu mu?"))]
oalErkekAnne1 <- nrow(filter(oal, CINSIYET == 1 & `Anne en fazla ilkokul mezunu mu?` == 1))
oalErkekAnne0 <- nrow(filter(oal, CINSIYET == 1 & `Anne en fazla ilkokul mezunu mu?` == 0))
oalErkekBaba1 <- nrow(filter(oal, CINSIYET == 1 & `Baba en fazla ilkokul mezunu mu?` == 1))
oalErkekBaba0 <- nrow(filter(oal, CINSIYET == 1 & `Baba en fazla ilkokul mezunu mu?` == 0))

oalKızAnne1 <- nrow(filter(oal, CINSIYET == 0 & `Anne en fazla ilkokul mezunu mu?` == 1))
oalKızAnne0 <- nrow(filter(oal, CINSIYET == 0 & `Anne en fazla ilkokul mezunu mu?` == 0))
oalKızBaba1 <- nrow(filter(oal, CINSIYET == 0 & `Baba en fazla ilkokul mezunu mu?` == 1))
oalKızBaba0 <- nrow(filter(oal, CINSIYET == 0 & `Baba en fazla ilkokul mezunu mu?` == 0))

xlab <- c("Mother Graduated", "Not", "Father Graduated", "Not")
ylab <- c(oalErkekAnne1, oalErkekAnne0, oalErkekBaba1, oalErkekBaba0,
          oalKızAnne1, oalKızAnne0, oalKızBaba1, oalKızBaba0)

oalMatrix <- matrix(ylab, 2, 4, byrow = TRUE, dimnames = list(c("Boy", "Girl"), xlab))

barplot(oalMatrix, col = c("red", "blue"), main = "Most Elementary School Graduations of OAL Students' Parents", names.arg = xlab)
```

```{r}
#Bar graph showing the most elementary school graduations of YEAL students' parents
yeal <- filter(df, OKUL == "YEAL")
yeal <- yeal[, (names(yeal) %in% c("CINSIYET", "Anne en fazla ilkokul mezunu mu?", "Baba en fazla ilkokul mezunu mu?"))]
yealErkekAnne1 <- nrow(filter(yeal, CINSIYET == 1 & `Anne en fazla ilkokul mezunu mu?` == 1))
yealErkekAnne0 <- nrow(filter(yeal, CINSIYET == 1 & `Anne en fazla ilkokul mezunu mu?` == 0))
yealErkekBaba1 <- nrow(filter(oal, CINSIYET == 1 & `Baba en fazla ilkokul mezunu mu?` == 1))
yealErkekBaba0 <- nrow(filter(oal, CINSIYET == 1 & `Baba en fazla ilkokul mezunu mu?` == 0))

yealKızAnne1 <- nrow(filter(oal, CINSIYET == 0 & `Anne en fazla ilkokul mezunu mu?` == 1))
yealKızAnne0 <- nrow(filter(oal, CINSIYET == 0 & `Anne en fazla ilkokul mezunu mu?` == 0))
yealKızBaba1 <- nrow(filter(oal, CINSIYET == 0 & `Baba en fazla ilkokul mezunu mu?` == 1))
yealKızBaba0 <- nrow(filter(oal, CINSIYET == 0 & `Baba en fazla ilkokul mezunu mu?` == 0))

ylabYeal <- c(yealErkekAnne1, yealErkekAnne0, yealErkekBaba1, yealErkekBaba0,
              yealKızAnne1, yealKızAnne0, yealKızBaba1, yealKızBaba0)

yealMatrix <- matrix(ylabYeal, 2, 4, byrow = TRUE, dimnames = list(c("Boy", "Girl"), xlab))

barplot(yealMatrix, col=c("red", "blue"),main="Most Elementary School Graduations of YEAL Students' Parents",names.arg=xlab)
```

```{r}
#qqplots
qqnorm(df$ORTALAMA, pch = 1, frame = FALSE,xlab = "GPA")
#qqline(dFile$ORTALAMA, col = "steelblue", lwd = 2)

#boxplots
boxplot(df$ORTALAMA,main="BOX PLOT FOR GPA")
```

```{r}
#Above average students
ort <- mean(df$ORTALAMA)
df <- transform(df, Başarılı.Mı= ifelse(ort >= ORTALAMA, 0, 1))
```

## Q1 Are students living in the village more successful than those living in the city center?
# Answer 1

```{r}
# Number of successful students (1) living in Village (0)
koyDf <- filter(df, IKAMET == 0)
basariliDf <- filter(df, Başarılı.Mı == 1)
koyIntersectDf <- intersect(koyDf, basariliDf)
koyBasariliSayisi <- nrow(koyIntersectDf)

# Number of successful students (1) living in the center (1)
merkezDf <- filter(df, IKAMET == 1)
merkezIntersectDf <- intersect(merkezDf, basariliDf)
merkezBasariliSayisi <- nrow(merkezIntersectDf)

if (koyBasariliSayisi > merkezBasariliSayisi) {
  print("Students living in the village are more successful.")
} else {
  print("Students living in the city center are more successful.")
}
  
```

## Q2 Are girls or boys most negatively affected by the separation of parents ?
# Answer 2

```{r}
#Mother and Father separate (1) number of girls (0) and unsuccessful (0) students
anneVeBabaAyriMiKızBasarisizDf <- filter(df, Anne.ve.babası.ayrı.mı. == 1 & CINSIYET == 0 & Başarılı.Mı == 0)
anneVeBabaAyriMiKızBasarisizSayisi <- nrow(anneVeBabaAyriMiKızBasarisizDf)

#Mother and Father separate (1) number of boys (0) and unsuccessful (0) students
anneVeBabaAyriMiErkekBasarisizDf <- filter(df, Anne.ve.babası.ayrı.mı. == 1 & CINSIYET == 1 & Başarılı.Mı == 0)
anneVeBabaAyriMiErkekBasarisizSayisi <- nrow(anneVeBabaAyriMiErkekBasarisizDf)

if (anneVeBabaAyriMiKızBasarisizSayisi > anneVeBabaAyriMiErkekBasarisizSayisi) {
  print("The students most affected by the separation of mothers and fathers are girl students.")
} else if (anneVeBabaAyriMiKızBasarisizSayisi < anneVeBabaAyriMiErkekBasarisizSayisi) {
  print("The students most affected by the separation of mothers and fathers are boy students.")
} else {
  print("Both genders were equally affected by the separation of the mother and father.")
}
```
## Q3 Which attribute is more important to success?
# Answer 3
```{r}
dfSorular <- df[, !(names(df) %in% c("OKUL", "SINIF", "NO", "ADI.SOYADI", "CINSIYET", "IKAMET", "ORTALAMA"))]

#install.packages("FSelectorRcpp")
library("FSelectorRcpp")
informatinGain <- information_gain(formula(dfSorular), dfSorular, equal = TRUE)
maxGain <- max(informatinGain$importance)
maxGainIndex <- match(maxGain, informatinGain$importance)

sprintf("The most important attribute for success : %s",informatinGain$attributes[maxGainIndex])

```

## Q4 Can a new student get a passing grade?
# Answer 4

```{r}
#https://www.geeksforgeeks.org/naive-bayes-classifier-in-r-programming/
#install.packages("e1071")
#install.packages("caTools")
#install.packages("caret")

library(ggplot2)
library(lattice)
library(e1071)
library(caTools)
library(caret)

# Splitting data into train
# and test data
split <- sample.split(dfSorular, SplitRatio = 0.7)
trainCl <- subset(dfSorular, split == "TRUE")
testCl <- subset(dfSorular, split == "FALSE")

# Feature Scaling
trainScale <- scale(trainCl[, 1:4])
testScale <- scale(testCl[, 1:4])

# Fitting Naive Bayes Model
# to training dataset
set.seed(120) #Setting Seed
classifierCl <- naiveBayes(formula(dfSorular), data = trainCl)

# Predicting on test data'
yPred <- predict(classifierCl, newdata = testCl)

# Confusion Matrix
cm <- table(testCl$Başarılı.Mı, yPred)

# Model Evaluation
confusionMatrix(cm)
```

## Q5 At which grade level is the educational status of the parents more effective?
# Answer 5

```{r}
edusta <- subset(df, df$Anne.en.fazla.ilkokul.mezunu.mu.==1 | df$Baba.en.fazla.ilkokul.mezunu.mu.==1)

edusta.dt9 <- data.table(edusta)[SINIF %like% "9"]
edusta.dt10 <- data.table(edusta)[SINIF %like% "10"]
edusta.dt11 <- data.table(edusta)[SINIF %like% "11"]
edusta.dt12 <- data.table(edusta)[SINIF %like% "12"]

edusta.dt9.mean <- mean(edusta.dt9$ORTALAMA)
edusta.dt10.mean <- mean(edusta.dt10$ORTALAMA)
edusta.dt11.mean <- mean(edusta.dt11$ORTALAMA)
edusta.dt12.mean <- mean(edusta.dt12$ORTALAMA)

A<-matrix(c(edusta.dt9.mean, edusta.dt10.mean, edusta.dt11.mean, edusta.dt12.mean), ncol = 1)
i <- 9
mingrade <- A[1,1]
for (val in A[ ,1]) {
  if(val<mingrade){
    mingrade <- val
    cls <-i
  } 
  i <- i+1
}
sprintf("The education level of the parents affected the %dth grades the most.",cls)

```