---
title: "APPLIED MACHINE LEARNING FINAL PROJECT"
subtitle: Hakkıcan BÜLÜÇ - 180709061 / Hüseyin ÇOBAN - 180709018 / Adem VAROL - 200709078
output:
  html_document: default
  pdf_document: default
font-family: Gill Sans
# author: "Hakkıcan BÜLÜÇ - 180709061 / Hüseyin ÇOBAN - 180709018 / Adem VAROL - 200709078"
# date: '2022-06-02'
date: "`r format(Sys.time(), '%d %B, %Y')`"
---
GROUP MEMBERS
Hakkıcan BÜLÜÇ - 180709061
Hüseyin ÇOBAN - 180709018
Adem VAROL - 200709078

Data Summary

Data read from excel file
```{r}
#install.packages("readxl")
library(readxl)
# Enter the path where the data is
df <- read_excel("01.AML_DATA_Etiketleri_Duzenlenmis.xls")
df
names(df)

summary(df$ORTALAMA)
```
```{r}
#install.packages("dplyr")
library(dplyr)
# Okul
# Ortaca Anadolu Lisesi (oal) student count
nrow(filter(df, OKUL == "OAL"))
```
```{r}
# Yunus Emre Anadolu Lisesi (yeal) student count
nrow(filter(df, OKUL == "YEAL"))
```
```{r}
# Class
library(data.table)
#df1$SINIF
dt9 <- data.table(df)[SINIF %like% "9"]
dt10 <- data.table(df)[SINIF %like% "10"]
dt11 <- data.table(df)[SINIF %like% "11"]
dt12 <- data.table(df)[SINIF %like% "12"]
# 9. grade count
nrow(dt9)
```
```{r}
# 10. grade count
nrow(dt10)
```
```{r}
# 11. grade count
nrow(dt11)
```
```{r}
# 12. grade count
nrow(dt12)
```
```{r}
# Sex
# boy
nrow(filter(df, CINSIYET == 1))
```
```{r}
# girl
nrow(filter(df, CINSIYET == 0))
```
```{r}
# From
# City Center
nrow(filter(df, IKAMET == 1))
```
```{r}
# Village
nrow(filter(df, IKAMET == 0))
```
```{r}
# Mean
mean(df$ORTALAMA)
```
```{r}
# Oal Mean
mean(filter(df, OKUL == "OAL")$ORTALAMA)
```
```{r}
# Yeal Mean
mean(filter(df, OKUL == "YEAL")$ORTALAMA)
```
```{r}
#Most successful student

#maxOrt <- max(df$ORTALAMA)
#maxOrtIndex <- match(maxOrt, df$ORTALAMA)
#df$ADI.SOYADI[maxOrtIndex]

df$`ADI SOYADI`[match(max(df$ORTALAMA), df$ORTALAMA)]

```
```{r}
#Bar graph showing the most elementary school graduations of Oal students' parents

oal <- filter(df, OKUL == "OAL")
oal <- oal[, (names(oal) %in% c("CINSIYET", "Anne en fazla ilkokul mezunu mu?", "Baba en fazla ilkokul mezunu mu?"))]
oalErkekAnne1 <- nrow(filter(oal, CINSIYET == 1 & `Anne en fazla ilkokul mezunu mu?` == 1))
oalErkekAnne0 <- nrow(filter(oal, CINSIYET == 1 & `Anne en fazla ilkokul mezunu mu?` == 0))
oalErkekBaba1 <- nrow(filter(oal, CINSIYET == 1 & `Baba en fazla ilkokul mezunu mu?` == 1))
oalErkekBaba0 <- nrow(filter(oal, CINSIYET == 1 & `Baba en fazla ilkokul mezunu mu?` == 0))

oalKızAnne1 <- nrow(filter(oal, CINSIYET == 0 & `Anne en fazla ilkokul mezunu mu?` == 1))
oalKızAnne0 <- nrow(filter(oal, CINSIYET == 0 & `Anne en fazla ilkokul mezunu mu?` == 0))
oalKızBaba1 <- nrow(filter(oal, CINSIYET == 0 & `Baba en fazla ilkokul mezunu mu?` == 1))
oalKızBaba0 <- nrow(filter(oal, CINSIYET == 0 & `Baba en fazla ilkokul mezunu mu?` == 0))

xlab <- c("Mother graduated", "not", "Father graduated", "not")
ylab <- c(oalErkekAnne1, oalErkekAnne0, oalErkekBaba1, oalErkekBaba0,
          oalKızAnne1, oalKızAnne0, oalKızBaba1, oalKızBaba0)

oalMatrix <- matrix(ylab, 2, 4, byrow = TRUE, dimnames = list(c("Boy", "Girl"), xlab))

barplot(oalMatrix, col = c("red", "blue"), main = "Showing The Most Elementary School Graduations of Oal Students' Parents", names.arg = xlab)
```

```{r}
#Bar graph showing the most elementary school graduations of Yeal students' parents

yeal <- filter(df, OKUL == "YEAL")
yeal <- yeal[, (names(yeal) %in% c("CINSIYET", "Anne en fazla ilkokul mezunu mu?", "Baba en fazla ilkokul mezunu mu?"))]
yealErkekAnne1 <- nrow(filter(yeal, CINSIYET == 1 & `Anne en fazla ilkokul mezunu mu?` == 1))
yealErkekAnne0 <- nrow(filter(yeal, CINSIYET == 1 & `Anne en fazla ilkokul mezunu mu?` == 0))
yealErkekBaba1 <- nrow(filter(oal, CINSIYET == 1 & `Baba en fazla ilkokul mezunu mu?` == 1))
yealErkekBaba0 <- nrow(filter(oal, CINSIYET == 1 & `Baba en fazla ilkokul mezunu mu?` == 0))

yealKızAnne1 <- nrow(filter(oal, CINSIYET == 0 & `Anne en fazla ilkokul mezunu mu?` == 1))
yealKızAnne0 <- nrow(filter(oal, CINSIYET == 0 & `Anne en fazla ilkokul mezunu mu?` == 0))
yealKızBaba1 <- nrow(filter(oal, CINSIYET == 0 & `Baba en fazla ilkokul mezunu mu?` == 1))
yealKızBaba0 <- nrow(filter(oal, CINSIYET == 0 & `Baba en fazla ilkokul mezunu mu?` == 0))

ylabYeal <- c(yealErkekAnne1, yealErkekAnne0, yealErkekBaba1, yealErkekBaba0,
              yealKızAnne1, yealKızAnne0, yealKızBaba1, yealKızBaba0)

yealMatrix <- matrix(ylabYeal, 2, 4, byrow = TRUE, dimnames = list(c("Boy", "Girl"), xlab))

barplot(yealMatrix, col = c("red", "blue"), main = "Showing The Most Elementary School Graduations of YEAL Students' Parents", names.arg = xlab)
```

```{r}
#qqplots
qqnorm(df$ORTALAMA, pch = 1, frame = FALSE,xlab = "ORTALAMA" )
#qqline(dFile$ORTALAMA, col = "steelblue", lwd = 2)

#boxplots
boxplot(df$ORTALAMA)
```

```{r}
#Calculate successful students

#df$ORTALAMA
ort <- mean(df$ORTALAMA)
#ort
df <- transform(df, Başarılı.Mı= ifelse(ort >= ORTALAMA, 0, 1))
#data.frame(df$ORTALAMA,df$Başarılı.Mı)
```

## Q1 Are students living in the village more successful than those living in the city center?

Answer 1

```{r}
# Number of successful students (1) living in Village (0)

#df$IKAMET
#df$Başarılı.Mı
koyDf <- filter(df, IKAMET == 0)
basariliDf <- filter(df, Başarılı.Mı == 1)
koyIntersectDf <- intersect(koyDf, basariliDf)
# Intersection df

#data.frame(koyIntersectDf$IKAMET, koyIntersectDf$Başarılı.Mı)

koyBasariliSayisi <- nrow(koyIntersectDf)

# Number of successful students (1) living in the center (1)
merkezDf <- filter(df, IKAMET == 1)
merkezIntersectDf <- intersect(merkezDf, basariliDf)
# Intersection df

#data.frame(merkezIntersectDf$IKAMET, merkezIntersectDf$Başarılı.Mı)

merkezBasariliSayisi <- nrow(merkezIntersectDf)

if (koyBasariliSayisi > merkezBasariliSayisi) {
  print("Students living in the village are more successful.")
} else {
  print("Students living in the city center are more successful.")
}
  
```

## Q2 Are girls or boys most negatively affected by the separation of parents ?

Answer 2

```{r}
#df$Anne.ve.babası.ayrı.mı.

#Mother and Father separate (1) number of girls (0) and unsuccessful (0) students

#df$CINSIYET
#df$Başarılı.Mı
anneVeBabaAyriMiKızBasarisizDf <- filter(df, Anne.ve.babası.ayrı.mı. == 1 & CINSIYET == 0 & Başarılı.Mı == 0)

anneVeBabaAyriMiKızBasarisizSayisi <- nrow(anneVeBabaAyriMiKızBasarisizDf)

#Mother and Father separate (1) number of boys (0) and unsuccessful (0) students
anneVeBabaAyriMiErkekBasarisizDf <- filter(df, Anne.ve.babası.ayrı.mı. == 1 & CINSIYET == 1 & Başarılı.Mı == 0)

anneVeBabaAyriMiErkekBasarisizSayisi <- nrow(anneVeBabaAyriMiErkekBasarisizDf)

if (anneVeBabaAyriMiKızBasarisizSayisi > anneVeBabaAyriMiErkekBasarisizSayisi) {
  print("The students most affected by the separation of mothers and fathers are girl students.")
} else if (anneVeBabaAyriMiKızBasarisizSayisi < anneVeBabaAyriMiErkekBasarisizSayisi) {
  print("The students most affected by the separation of mothers and fathers are boy students.")
} else {
  print("Both genders were equally affected by the separation of the mother and father.")
}
```

## Q3 Which attribute is more important to success?

Answer 3

```{r}
#to remove a Column

#df = df[,!(names(df) %in% c("Geçer.Mİ","Geçer.Mİ.1", "Geçer.Mİ.2"))]

#we only got questions from df

#names(df)
dfSorular <- df[, !(names(df) %in% c("OKUL", "SINIF", "NO", "ADI.SOYADI", "CINSIYET", "IKAMET", "ORTALAMA"))]

#install.packages("FSelectorRcpp")
library("FSelectorRcpp")
informatinGain <- information_gain(formula(dfSorular), dfSorular, equal = TRUE)
maxGain <- max(informatinGain$importance)
maxGainIndex <- match(maxGain, informatinGain$importance)
informatinGain$attributes[maxGainIndex]
```

## Q4 Can a new student get a passing grade?

Answer 4

```{r}
#https://www.geeksforgeeks.org/naive-bayes-classifier-in-r-programming/
#install.packages("e1071")
#install.packages("caTools")
#install.packages("caret")

library(ggplot2)
library(lattice)
library(e1071)
library(caTools)
library(caret)

# Splitting data into train
# and test data
split <- sample.split(dfSorular, SplitRatio = 0.7)
trainCl <- subset(dfSorular, split == "TRUE")
testCl <- subset(dfSorular, split == "FALSE")

# Feature Scaling
trainScale <- scale(trainCl[, 1:4])
testScale <- scale(testCl[, 1:4])

# Fitting Naive Bayes Model
# to training dataset
set.seed(120) #Setting Seed
classifierCl <- naiveBayes(formula(dfSorular), data = trainCl)

# Predicting on test data'
yPred <- predict(classifierCl, newdata = testCl)

# Confusion Matrix
cm <- table(testCl$Başarılı.Mı, yPred)

# Model Evaluation
confusionMatrix(cm)
```

## Q5 At which grade level is the educational status of the parents more effective?

Answer 5

```{r}
#df$Baba.en.fazla.ilkokul.mezunu.mu.
#df$`Baba en fazla ilkokul mezunu mu?`
df1 <- transform(df, Anne.ve.Baba.en.fazla.ilkokul.mezunu.Mu = ifelse(Anne.en.fazla.ilkokul.mezunu.mu. & Baba.en.fazla.ilkokul.mezunu.mu., 1, 0))

#data.frame(df1$Anne.en.fazla.ilkokul.mezunu.mu., df1$Baba.en.fazla.ilkokul.mezunu.mu., df1$Anne.ve.Baba.en.fazla.ilkokul.mezunu.Mu)
dt9 <- data.table(df1)[SINIF %like% "9"]
dt10 <- data.table(df1)[SINIF %like% "10"]
dt11 <- data.table(df1)[SINIF %like% "11"]
dt12 <- data.table(df1)[SINIF %like% "12"]

#dt9$Anne.ve.Baba.en.fazla.ilkokul.mezunu.Mu
dt9 <- filter(dt9, `Anne.ve.Baba.en.fazla.ilkokul.mezunu.Mu` == 0)
dt10 <- filter(dt10, `Anne.ve.Baba.en.fazla.ilkokul.mezunu.Mu` == 0)
dt11 <- filter(dt11, `Anne.ve.Baba.en.fazla.ilkokul.mezunu.Mu` == 0)
dt12 <- filter(dt12, `Anne.ve.Baba.en.fazla.ilkokul.mezunu.Mu` == 0)

dt9Mean <- mean(dt9$ORTALAMA)
dt10Mean <- mean(dt10$ORTALAMA)
dt11Mean <- mean(dt11$ORTALAMA)
dt12Mean <- mean(dt12$ORTALAMA)

maxSinif <- max(dt9Mean, dt10Mean, dt11Mean, dt12Mean)
#maxSinif

if (dt9Mean > dt10Mean && dt9Mean > dt11Mean && dt9Mean > dt12Mean) {
  print("The education level of the parents affected the 9th grades the most.")
} else if (dt10Mean > dt9Mean && dt10Mean > dt11Mean && dt10Mean > dt12Mean) {
  print("The education level of the parents affected the 10th grades the most.")
} else if (dt11Mean > dt9Mean && dt11Mean > dt10Mean && dt11Mean > dt12Mean) {
  print("The education level of the parents affected the 11th grades the most.")
} else {
  print("The education level of the parents affected the 12th grades the most.")
}
```